### 2020-02-17 chapters/appendix1.tex

\chapter{Loren Ipsum} \label{ap1:loren}

If you are going to use a passage of Lorem Ipsum, you need to be sure
there isn't anything embarrassing hidden in the middle of text.

\section{What is \emph{Loren Ipsum}?}

\emph{\textbf{Lorem Ipsum}} is simply dummy text of the printing and
typesetting industry. Lorem Ipsum has been the industry's standard
dummy text ever since the 1500s, when an unknown printer took a galley
of type and scrambled it to make a type specimen book.
It has survived not only five centuries, but also the leap into
electronic typesetting, remaining essentially unchanged.
It was popularised in the 1960s with the release of Letraset sheets
containing Lorem Ipsum passages, and more recently with desktop
publishing software like Aldus PageMaker including versions of Lorem
Ipsum\footnote{Available at \url{http://www.lipsum.com/}}.

\section{Where does Loren come from?}

Contrary to popular belief, Lorem Ipsum is not simply random text.
It has roots in a piece of classical Latin literature from 45 BC,
making it over 2000 years old.
Richard McClintock, a Latin professor at Hampden-Sydney College in
Virginia, looked up one of the more obscure Latin words, consectetur,
from a Lorem Ipsum passage, and going through the cites of the word in
classical literature, discovered the undoubtable source.
Lorem Ipsum comes from sections 1.10.32 and 1.10.33 of ``de Finibus
Bonorum et Malorum'' (The Extremes of Good and Evil) by Cicero,
written in 45 BC.
This book is a treatise on the theory of ethics, very popular during
the Renaissance. T
he first line of Lorem Ipsum, ``Lorem ipsum dolor sit amet\ldots'',
comes from a line in section 1.10.32.

The standard chunk of Lorem Ipsum used since the 1500s is reproduced
below for those interested.
Sections 1.10.32 and 1.10.33 from ``de Finibus Bonorum et Malorum'' by
Cicero are also reproduced in their exact original form, accompanied
by English versions from the 1914 translation by H. Rackham.

\section{Why using Loren?}

It is a long established fact that a reader will be distracted by the
readable content of a page when looking at its layout.
The point of using Lorem Ipsum is that it has a more-or-less normal
distribution of letters, as opposed to using ``Content here, content
here'', making it look like readable English.
Many desktop publishing packages and web page editors now use Lorem
Ipsum as their default model text, and a search for ``lorem ipsum''
will uncover many web sites still in their infancy.
Various versions have evolved over the years, sometimes by accident,
sometimes on purpose (injected humour and the like).

\section{Where to Find Examples?}

There are many variations of passages of Lorem Ipsum available, but
the majority have suffered alteration in some form, by injected
humour, or randomised words which don't look even slightly
believable.
If you are going to use a passage of Lorem Ipsum, you need to be sure
there isn't anything embarrassing hidden in the middle of text.
All the Lorem Ipsum generators on the Internet tend to repeat
predefined chunks as necessary, making this the first true generator 
on the Internet.
It uses a dictionary of over 200 Latin words, combined with a handful
of model sentence structures, to generate Lorem Ipsum which looks
reasonable.
The generated Lorem Ipsum is therefore always free from repetition,
injected humour, or non-characteristic words etc.