### 2020-02-17 chapters/conclusion.tex

\chapter{Conclusion} \label{chap:conclusion} \minitoc

We introduced background concepts on large scale data processing, windowing techniques applied to data streams and sliding window aggregations. We studied algorithms for change detection in data streams, pattern mining, anomaly detection and sliding window aggregation algorithms, specifically a subset of these named probabilistic data structures.

%%hypothesis validated?
The monitoring of data streams in real-time is applicable across many domains and applications. To the best of our knowledge, we claim that there has not been a real attempt at doing so in a lightweight way.

%%future work
Preliminary work shows that there exist constant space and time complexity algorithms to aggregate data, thus reducing memory usage. We intend to research more about lightweight sliding window aggregations and sliding implementations of probabilistic data structures. We also intend to search for new methods of summarizing data that do not necessarily explicitly work in a sliding window fashion. After identifying a set of algorithms that meet our requirements, the focus will be on implementing the best theoretical solution and benchmark our streaming system using data sets from the financial fraud space versus previously performed batch analysis.

%%summary of contributions
The main contribution of this Thesis will be the developed system code as well as a set of tests that measure the system's usefulness based on several parameters, such as recall, precision, memory used and latency.