### 2020-03-06 chapters/conclusion.tex
Aggregation on data types

\chapter{Conclusion} \label{chap:conclusion} \minitoc

Yeeeeeeetttttt