### 2020-07-02 chapters/conclusions.tex
Restivo feedback applied

\chapter{Conclusion} \label{chap:conclusion} \minitoc

\section{Hypothesis summary}
\section{Answer to the hypothesis}
\section{Contributions}
\section{Future work}
\section{Final remarks}