### 2020-07-02 chapters/conclusions.tex
Capitalization

\chapter{Conclusion} \label{chap:conclusion} \minitoc

\section{Hypothesis Summary}
\section{Answer to the Hypothesis}
\section{Contributions}
ema like hist

lightweight real-time method

set of synthetic test

experiments on real data

\section{Future Work}

- debug it

- test it

- wassertein

- other then holm

- other then jsd

\section{Final Remarks}